---
title   : "1 - Introduction to R"
author  : "Yinan Guo"
date    : "July 9th, 2021" 
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: paper
    highlight: tango
---

## Instructions 

Lets jump right into R coding this week! In this assignment, we'll be creating a "Cheat Sheet" that you can refer back to over the course of this section and we'll jump into some basic analysis with ggplot2. To complete this assignment, follow these steps:

1. Download the `introduction_to_R_yournamehere.Rmd` file from the course website.

2. Open `introduction_to_R_yournamehere.Rmd` in RStudio.

3. Replace the "Your Name Here" text in the `author:` field with your own name.

4. Supply your solutions to the project by editing `introduction_to_R_yournamehere.Rmd`.

5. When you have completed the homework and have **checked** that your code both runs in the Console and **knit** correctly when you click `Knit HTML`, save the Knitted HTML file as `introduction_to_R_yournamehere.html`, and submit the HTML file!!  

> Tip:  Note that each of the code blocks in this Problem contain the expression `eval = FALSE`.  This tells R Markdown to display the code contained in the block, but not to evaluate it.  you'll need to change this to `eval = TRUE` before you

**knit**

### Problem 0. Load Libraries 
> load the following libraries 
- tidyverse
- skimr
- janitor 

don't forget to change `eval=FASE` to `eval=TRUE`

```{r}
library(tidyverse)
library(skimr)
library(janitor)
```


### Problem 1: Data Types 

> In this section we'll work with R's two base data types Numeric & Character as well Integer, Logial and Dates to do this answer the following questoins 


##### (a) Numeric Data using R as a Calculator 

In its most basic form, R can be used as a simple calculator using basic arithmetic operators:

- Addition: +
- Subtraction: -
- Multiplication: *
- Division: /
- Exponentiation: ^
- Modulo: %%

> question 1. write an expression that evaluates to TRUE if x == y ^ z where you assign x to 128 and assign y to 2 what does z need to be? 

don't forget to change `eval=FASE` 

```{r}
  # Insert your Assignment logic here
x <- 128
y <- 2
z <- log(x,y)

  # Insert your Boolean Expression here 
x == y^z

```


##### (b) Checking inequality.

> question 2. write an equation that evaluates to TRUE if x / y is greater than 1/x times y.

```{r}
x <- 5
y <- 4

# Insert your Boolean expression here
x / y >= 1/x * y 
```

##### (c) A formula. 

> question 3. given the equation IBU = -32.9 + 1156 * abv + 25.27 * ipa + 2.79 * ale 

If you replace abv with 0.062, ipa with 0 and ale with a 1 what does ibu sum up to? 

Is ibuthan 50 or less than 50?  

hint: there may be a if/else gotcha down below... 


```{r}
abv <- 0.062
ipa <- 0
ale <- 1
ibu = -32.9 + 1156 * abv + 25.27 * ipa + 2.79 * ale

L <- ibu > 50

# Insert your Boolean expression here
if (L ){
  print("Greater than 50  ")
} else {
  print("Less than 50  ")
}
```


##### (d) playing with character data.

> question 4. write an expression using sprintf to print out "George Washington Hayduke has $250.01" given the following 

```{r}
x <- "Seldom Seen Smith"

fname <- "George"
mname <- "Washington"
lname <- "Hayduke"
money <- 250.01 

full_name <- paste(fname, mname, lname, sep=" " )

sprintf("%s %s%.2f",full_name, "has $", money)
```


### Problem 2: Vectors 

#### (a) create a numeric vector 

> question 1. craete a vector containing of 20 random numbers from 0 to 100 and peform the following operations on it

- use runif to create your vector ranif(20, min=0, max=100)
- use vector arithmetic so that the vector sums to 1  

```{r, eval = TRUE}
v1 <- runif(20, min=0, max=100)

# use vector arithmetic so that the v2 sums to 1, simply v1/sum(v1)
v2 <- v1 / sum(v1)
sum(v2)
```



##### (b) Comparing  Character Vectors 

> Given two users's preferences between two options: [Facebook, Instagram], [Firefox, Chrome], [Mac, PC], [Summer, Winter].  
- how many things do they aggree on and how many do they not? 


```{r, eval = TRUE}
mike  <- c("Instagram", "Chrome", "Mac", "Summer")
cole  <- c("Facebook", "Chrome", "PC", "Winter")

same_n <- length(intersect(mike,cole))
same_n

diff_n <- length(setdiff(mike,cole))
diff_n
```

##### (c) Looping through a vector  

> use a for loop to itterate through a list of numbers  

for (num in list_o_numbers){
  print(num)
}

```{r, eval = TRUE}
list_o_numbers <- runif(20, min=0, max=100)

for (num in list_o_numbers){
  print(num)
}
```

### Problem 3: Data frame basics
> grab the AUTO_MPG.csv from the course website 

##### (a) Importing data.

> I prefer to use the `readr` libary for reading CSV files into R as apposed to the base `read.csv()` function to import data. use both readr's `read_csv` and `read.csv()` 
- use head to disply the first 5 records of each. 
- do you notice anything different? 

**Tip**: if the data is located at a URL you can import it simply by referencing the URL you don't need to always download the data first,  instead you can import it  data directly using the URL.

```{r, eval=TRUE}
library(readr)
mpg_1 <- read_csv('AUTO_MPG.csv')

mpg_2 <- read.csv('AUTO_MPG.csv')

# display the top 5 records of both using head(mpg_1) and head(mpg_2)... what do you notice different? 
head(mpg_1)

head(mpg_2)

```
read_csv will generate data into tbl_df.
read.csv will show the number of row.