---
title   : "Assignment 2 - Austin TX Coffee Sentiment"
author  : "Yinan Guo"
date    : "2021-07-18" 
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: paper
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

The Ratings_and_Sentiments.csv dataset contains over 7,600 reviews of 66 coffee shops in Austin Texas, the data was initially scraped, munged and prepped by Rachel Downs (http://www.racheldowns.co) a Marketing and MIS student at UT Austin. 

Your challenge is to further prepare the data (easy), and create summaries and charts answering various questions about the data (also easy) 

To complete this assignment, follow these steps:

1. Download the `2_cup_o_joe_yournamehere.RMD` file from the course website.

2. Open `2_cup_o_joe_yournamehere.RMD` in RStudio.

3. Replace the "Your Name Here" text in the `author:` field with your own name.

4. Supply your solutions to the project by editing `2_cup_o_joe_yournamehere.RMD`.

5. When you have completed the homework and have **checked** that your code both runs in the Console and **knit** correctly when you click `Knit HTML`, rename the R Markdown file to `Austin_Coffee_Sentiment_YourNameHere.Rmd`, and submit it.  

> Tip:  Note that each of the code blocks in this Problem contain the expression `eval = FALSE`.  This tells R Markdown to display the code contained in the block, but not to evaluate it.  you'll need to change this to `eval = TRUE` before you **knit!!!**

### Step 0. Load Libraries 

> load the following libraries, you may need to install them first! 
- tidyverse 
- lubridate # -- a new library that makes dealing with dates easy 
- stringr   # -- a new library that makes dealing wiht strings easy 
- ggplot2
- janitor 
- skimr 

don't forget to change `eval=FASE` to `eval=TRUE`

```{r}
library(tidyverse)
library(skimr)
library(janitor)
library(stringr) 
library(lubridate)
library(ggplot2)
library(kableExtra)
```


### Step 1. Stage 

Read (read_csv), ignore warnings - you'll see 4 parsing failures just ignore them. 

1. create a new data frame called "yelp" using read_csv() to read in ratings_and_sentiments.csv data file 

2. display the first 5 records

```{r}
yelp <-read_csv("data/ratings_and_sentiments.csv")
head(yelp,5) %>%
  kable() %>%
  kable_styling(font=8)
```


### Step 2. Structure & Transform 

Using the PIPE (%>%) and stringr functions  to create a "yelp_prep" dataset  by performing the following: 

1.  Filter out any rows where coffee_shop_name is Null: filter(!is.na(coffee_shop_name))

2.  Create a new date variable (review_date) based on the date contained the review_text field using mutate.

- You'll note the review date is the first item in that field, so you'll need to parse it. 
- if you've loaded stringr the "word()" function will grab the first "word" in a sentance 
- or you can do the hard route  something like this ~  sub("\\s.*","", review_text)
- To convert review_date to a Date data type you'll need to use another function 
- lubridate has a function called mdy() for month day and year it will convert a string into a date, use that! 
- make sure your new field is a DATE data type

> Hint: mutate(review_date = mdy(word(review_text)))

3. Create a flag variable "starbucks_flag", if the review mentions "starbucks", regardless of case, set the flag to "yes" else default it to "no" 

> Hint: use mutate() to create starbucks_flag, use if_else(condition,1,0), there should be at least 320 mentions of starbucks 
https://stringr.tidyverse.org/reference/index.html
https://stringr.tidyverse.org/reference/str_detect.html

> Hint: stringr and and str_detect() how do you deal with mixed cases hmm?
https://stringr.tidyverse.org/reference/case.html


4. Create a flag variable "good_flag" if the review mentions the word "good", regardless of case, set the flag to "yes" else default it to "no" , there should be around 3077 good references 

5. Create a flag variable "great_flag" if the review mentions the word "great", regardless of case, set the flag to "yes" else default it to "no" , there should be around 2870 references to "great". 

6. Convert coffee_sent to numeric using as.numeric(coffee_sent) 

7. list the first 5 records to make sure your code works. 



```{r}
yelp_prep <- yelp %>%
  #Q1
  filter(!is.na(coffee_shop_name)) %>%
  
  #Q2
  mutate(review_date = mdy(word(review_text))) %>%
  
  #Q3
  mutate(starbucks_flag=if_else(str_detect(str_to_lower(review_text),"starbucks"),"yes","no")) %>%
  
  #Q4
  mutate(good_flag=if_else(str_detect(str_to_lower(review_text),"good"),"yes","no")) %>%
  
  #Q5
  mutate(great_flag=if_else(str_detect(str_to_lower(review_text),"great"),"yes","no")) %>%
  
  #Q6 (there is a #VALUE! data but not null in coffee_sent so I filter it out)
  filter(coffee_sent %in%  c("-3","-2","-1","0","1","2","3","4")) %>%
  mutate(coffee_sent=as.numeric(coffee_sent)) 
  
  #Q7
head(yelp_prep,5)%>%
  kable() %>%
  kable_styling(font=8)
```


### Step 3. Frequency Analysis & Graphs 

Create 5 bar charts, for the following variables, set the fill/color setting by indicated fields, use a facet wrap for the following. What if anything do the graphs tell you about sentiment and ratings? 

if the fill/color variable is a numeric be sure to change it to a factor! 

> here is a template to follow 

yelp_prep %>%
  ggplot(aes(x=numeric_variabale, fill=fill_variable)) +
  geom_bar() + 
  facet_wrap(~facet_variable) + 
  labs(title = , x = , y = )


Variable    Fill/Color  Facet Wrap      
----------  ----------  --------------
num_rating  great_flag  starbucks_flag


```{r}
yelp_prep %>%
  ggplot(aes(x=num_rating, fill=great_flag)) +
  geom_bar() + 
  facet_wrap(~starbucks_flag) + 
  labs(title = "Rating vs. 'STARBUCKS' Mentioned vs. 'GREAT' Mentioned",subtitle = "'STARBUCKS' Mentioned", x = "Rating Num", y = "Count of Ratings")
```


Variable      Fill/Color      facet wrap
------------  --------------  ----------
overall_sent  starbucks_flag  NA


```{r}
yelp_prep %>%
  ggplot(aes(x=overall_sent, fill=starbucks_flag)) +
  geom_bar() + 
  labs(title = "Overall Sentiment vs. 'STARBUCKS' Mentioned", x = "Overall Sentiment", y ="Count of Overall Sentiment" )
```


Variable       Fill/Color   Facet Wrap      
------------   ----------   --------------
service_sent    great_flag  


```{r}
yelp_prep %>%
  ggplot(aes(x=service_sent, fill=great_flag)) +
  geom_bar() + 
  labs(title = "Service Sentiment vs. 'GREAT' Mentioned", x = "Service Sentiment", y ="Count of Service Sentiment" )
```

Variable      Fill/Color  Facet Wrap      
------------  ----------  --------------
coffee_sent   N/A         Starbucks_flag



```{r}
yelp_prep %>%
  ggplot(aes(x=coffee_sent)) +
  geom_bar() + 
  facet_wrap(~starbucks_flag) + 
  labs(title = "Coffee Sentiment vs. 'STARBUCKS' Mentioned",  x = "Coffee Sentiment", y ="Count of Coffee Sentiment" )
```


Variable    Fill/Color  Facet Wrap      
----------  ----------  --------------
food_sent   good_flag   N/A


```{r}
yelp_prep %>%
  filter(!is.na(food_sent)) %>%
  #There is an NA data in food_sent so I filter it out.
  ggplot(aes(x=food_sent, fill=good_flag)) +
  geom_bar() + 
  labs(title = "Food Sentiment vs. 'GOOD' Mentioned", x = "Food Sentiment", y ="Count of Food Sentiment" )
```



### Step 4. More interesting Analysis, Does day of Week Matter? 

Create a new data set called sentiment_summary

1. Create a new varaible day_of_week by applying wday to review_date, like this: wday(review_date, label = TRUE, abbr = FALSE)
2. Group by day_of_week 
3. Use summarize to create a count, mean of overall_sent, a mean of coffee_sent, rememver to remove NA values with na.rm = TRUE

> also if you forgot to convert coffee_sent to numeric now is a good time to go back and do that : ) 


```{r}
yelp_final <- yelp_prep %>%
  mutate(day_of_week=wday(review_date, label = TRUE, abbr = FALSE)) %>%
  group_by(day_of_week) %>%
  summarise(n=n(), mean_overall_sent=mean(overall_sent,na.rm=TRUE), mean_coffee_sent=mean(coffee_sent,na.rm=TRUE))
yelp_final%>%
  kable() %>%
  kable_styling(font=10)
```


## Step 5. Create the Following Bar Charts using stat="identity"

Create three Bar charts to answer the following questions 

1. what is the most/least reviewed day of the week 
Sunday has the most and Friday has the least reviews.

2. what day are you most likely to get the highest/lowest mean overall sentiment 
Wednesday the highst and Sunday the lowest.

3. what day are you most likely to get the highest/lowest mean coffee sentiment 
Monday the highest and Wednesday the lowest.
  


```{r}
plot1 <-
  ggplot(yelp_final) +
  geom_bar(mapping=aes(x=day_of_week, y=n,fill=day_of_week), stat = "identity") + 
  labs(title = "REVIEWS FOR DAYS OF WEEK", x = "DAY OF WEEK", y = "REVIEW NUM" )
plot1

plot2 <-
  ggplot(yelp_final) +
  geom_bar(aes(x=day_of_week, y=mean_overall_sent,fill=day_of_week), stat = "identity") + 
  labs(title = "MEAN OF OVERALL SENTIMENT FOR DAYS OF WEEK", x = "DAY OF WEEK", y = "MEAN OF OVERALL SENTIMENT")
plot2

plot3 <-
  ggplot(yelp_final) +
  geom_bar(aes(x=day_of_week, y=mean_coffee_sent,fill=day_of_week), stat = "identity") + 
  labs(title = "MEAN OF COFFEE SENTIMENT FOR DAYS OF WEEK", x = "DAY OF WEEK", y = "MEAN OF COFFEE SENTIMENT")
plot3
```


## Finally

Wrap this up in a R notebook and knit it to HTML, to exceed the bar - make sure things look good. 

